---
title: "a4"
author: "Mushi Wang"
date: "27/11/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## q1
```{r}
library(readxl)
carb = read_excel("carbonation.xls")
carb
```
**(a)**
```{r}
fit = lm(y~., carb)
par(mfrow=c(2,2))

plot(fitted(fit), residuals(fit), xlab="fitted value", ylab="Residuals")

plot(carb$x1, residuals(fit), xlab="x1", ylab="Residuals")

plot(carb$x2, residuals(fit), xlab="x2", ylab="Residuals")
```
For the plots of fitted value vs residuals and $x_2$ vs residuals, there is a qudratic pattern. 
For the plot of $x_1$ vs residuals, the absoluate value of residuals of $x_1 = 31$ is greater than the other $x_1$ values.
Hence the fitted model is not adequate.

*(b)*
```{r}
fit2 = lm(y~poly(x1, 2) + poly(x2, 2), carb)
summary(fit2)
```
$x_2$ and ${x_2}^2$ are significant at significant level 0.05.
$x_1$ and ${x_1}^2$ are not significant.

*(c)*
```{r}
par(mfrow=c(2,2))

plot(fitted(fit2), residuals(fit2), xlab="fitted value", ylab="Residuals")

plot(carb$x1, residuals(fit2), xlab="x1", ylab="Residuals")

plot(carb$x2, residuals(fit2), xlab="x2", ylab="Residuals")
```
There are no systematic patterns in any plots, and the residuals lie within a band around 0.
Hence, the fitted model is adequate.

*(d)*
```{r}
qqnorm(residuals(fit2))
qqline(residuals(fit2))
```
The points in QQ plot look approxiamtely in a straight line.
Hence, the residual is normally distributed.

*(e)*
```{r}
summary(fit)
```
The adjusted $R^2$ of the model in part(b) is greater than the model in part(a), so we prefer the model in part(b).

##q2
```{r}
sal = read.table("salary.txt", header = FALSE)
colnames(sal) = c("y", "degree", "exp", "sup")
```

*(a)*
```{r}
deg = factor(sal$degree)
sallm = lm(y ~ deg + exp + sup, data = sal)
summary(sallm)
```
```{r}
plot(fitted(sallm), residuals(sallm), xlab = "fitted value", ylab = "residuals")
```
The residuals are fan shaped, the variances of random errors are non-constant.

*(b)*
```{r}
par(mfrow = c(1,2))
plot(sal$exp, abs(residuals(sallm)), xlab = "x3", ylab = "residuals")
plot(sal$sup, abs(residuals(sallm)), xlab = "x4", ylab = "residuals")
```
The residuals are fan shaped, the variances of random errors are non-constant.

*(c)*
```{r}
r = abs(residuals(sallm))
rsd = lm(r~sal$exp + sal$sup)
1/(fitted(rsd)^2)
```

*(d)*
```{r}
sal$wts = 1/(fitted(rsd)^2)
sallm2 = lm(y ~ deg + exp + sup, weights = wts, data = sal)
summary(sallm2)
```
These estimates are similar to the the estimates in part(a).

*(e)*
most of the deviations are less that part(a). So the second model is more precise.

*(f)*
```{r}
plot(fitted(sallm2), residuals(sallm2), xlab="fitted value", ylab="Residuals")
```
The residual plot is still fan shaped. Since the estimates from two models are relatively similar, we expect a similar residual plot.

```

