---
title: "a4"
author: "Mushi Wang"
date: "15/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***q1***
```{r}
options(contrasts = c('contr.sum', 'contr.poly'))

time1 = c(9, 12, 10, 8, 15)
time2 = c(20, 21, 23, 17, 30)
time3 = c(6, 5, 8, 16, 7)
y = c(time1, time2, time3)
type = as.factor(c(rep(1, 5), rep(2, 5), rep(3, 5)))

model = lm(y ~ type)
summary(model)
```
(a)
$H_0: \tau_1 = \tau_2 = \tau_3 = 0$ vs $H_a :$ at least one of them is not $0$.

From summary, F-statistic is $16.08$ on $2$ and $12$ DF. The p-value is $0.0004023$. So there is tons of evidence reject $H_0$. Hence, the three circuit types have different response time.

(b)
$\theta = \tau_2 - \frac {\tau_1 + \tau_3} {2}$, 
$\tilde\theta = \tilde\tau_2 - \frac {\tilde\tau_1 + \tilde\tau_3} {2}$

(c)
Form summary, $\hat\tau_1 = -3$, $\hat\tau_2 = 8.4$, and $\hat\tau_3 = -(-3 + 8.4) = -5.4$. Since, $\bar{y}_{i+} = \bar{y}_{++} + \tau_{i}$ and $Var(\bar{Y}_{1+}) = Var(\bar{Y}_{2+}) = Var(\bar{Y}_{3+})$. Therefore the widths of confidence intervals of $\bar{Y}_{1+}, \bar{Y}_{2+}, \bar{Y}_{3+}$ has the same width.So we choose circuit type 3.

(d)
```{r}
anova(model)
```
```{r}
par(mfrow=c(2,2))
qqnorm(model$residuals)
plot(model$residuals)
plot(model$fitted.values, model$residuals)
```
In Q-Q plot, all the points are lie in a straight line with some exceptions on the right. So there could be a violation against assumption.

In indices vs residuals, all points are lie in a band between 0 and there is no obvious pattern.

In fitted values vs residuals, all points are lie in a band between 0 and there is no obvious pattern. But the could be a violation against assumption since we only have three distinct fitted values.


\newpage
***q2***
```{r}
effectiveness = c(c(13, 22, 18, 39), c(16, 24, 17, 44), c(5, 4, 1, 22))
type = as.factor(c(rep(1, 4), rep(2, 4), rep(3, 4)))
block = as.factor(c(seq(from = 1, to = 4), seq(from = 1, to = 4), seq(from = 1, to = 4)))

effect = lm(effectiveness ~ type + block)
```

(a)
$H_0: \tau_1 = \tau_2 = \tau_3 = 0$ vs $H_a :$ at least one of them is not $0$
```{r}
anova(effect)
```

Form anova table, $f = 40.717$, and p-value $= 0.0003232$. So there is tons of evidence reject $H_0$. Hence solutions have different effectiveness.

(b)
$H_0: \beta_1 = \beta_2 = \beta_3 = \beta_4= 0$ vs $H_a :$ at least one of them is not $0$
```{r}
anova(effect)
```
Form anova table, $f = 42.711$, and p-value $= 0.00019252$. So there is tons of evidence reject $H_0$. Hence solutions have different effectiveness.

(c)
$H_0: \bar{y}_{2+} > 30$ vs $H_a: \bar{y}_{2+} \leq 30$

Let $\theta = \bar{Y}_{2+}$
```{r}
summary(effect)
```
\begin{align*}
    E(\tilde{\theta}) & = E(\tilde{\tau_{2}} + \bar{Y}) \\
    & = E(\tilde{\tau_{2}}) + E(\bar{Y}) \\
    & = \tau_2 + \mu \text{\quad since unbiased}
\end{align*}
and,
\begin{align*}
    Var(\tilde{\theta}) & = Var(\tilde{\tau_{2}} + \bar{Y}) \\
    & = Var(\bar{Y_2} - \bar{Y} + \bar{Y}) \\
    & = Var(\bar{Y_2}) \\
    & = \frac {\sigma^2} {4} 
\end{align*}
So, $\hat{\theta} = \hat{\mu} + \hat{\tau_2} = 18.75 + 6.5 = 25.25$
\begin{align*}
    d & = \frac {\hat{\theta} - 30} {se(\tilde{\theta})} \\
    & = \frac {25.25 - 30} {2.939 / \sqrt{4}} \\
    & = -3.232392
\end{align*}
Since $D \sim t_{12-7+2-1} = t_{6}$
```{r}
pt(-3.232392, 6)
```
p-value = $0.008928362$, so there is tons of evidence reject $H_0$. Hence the mean of solution 2 is less than or equal 30.

(d)
$H_0: \tau_1 - 2 \tau_2 + \tau_3 = 0$ vs $H_a: \tau_1 - 2 \tau_2 + \tau_3 \neq 0$
```{r}
summary(effect)
```
Let $\theta = \tau_1 - 2 \tau_2 + \tau_3$

\begin{align*}
    E(\tilde{\theta}) & = E(\tilde{\tau_1}) - 2 E(\tilde{\tau_2}) + E(\tilde{\tau_3}) \\
    & = \tau_1 - 2 \tau_2 + \tau_3 \text{\quad since unbiased}
\end{align*}
and, 
\begin{align*}
    Var(\tilde{\theta}) & = Var(\tilde{\tau_1} - 2 \tilde{\tau_2} + \tilde{\tau_3}) \\
    & = Var(\tilde{\tau_1}) + 4 Var(\tilde{\tau_2}) + Var(\tilde{\tau_3}) \\
    & = \frac {\sigma^2} {4} + \sigma^2 + \frac {\sigma^2} {4} \\
    & = \frac 3 2 \sigma^2
\end{align*}
So,
\begin{align*}
    d & = \frac {\hat{\tau_1} - 2 \hat{\tau_2} + \hat{\tau_3} - 0} {\sqrt{\frac 3 2 \hat{\sigma}^2}}\\
    & = \sqrt{2} \frac {4.25 - 2 \times 6.5 - (4.25 + 6.5)} {\sqrt{3} \times 2.939} \\
    & = -5.417381
\end{align*}

Since $D \sim t_{12-7+2-1} = t_{6}$
```{r}
2 * (1 - pt(5.417381, 6))
```
p-value = $2p(D > \lvert d\rvert)$ = 0.001636828

So we have tons of evidence to reject $H_0$. Hence $\tau_1 - 2 \tau_2 + \tau_3 \neq 0$
