---
title: "a3q1"
author: "Mushi Wang"
date: "11/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.
```{r}
options(contrasts = c('contr.sum', 'contr.copy'))
dia1 = c(80, 83, 83, 85)
dia2 = c(75, 75, 79, 79)
dia3 = c(74, 73, 76, 77)
dia4 = c(67, 72, 74, 74)
dia5 = c(62, 62, 67, 69)
dia6 = c(60, 61, 64, 66)
y = c(dia1, dia2, dia3, dia4, dia5, dia6)
x = as.factor(c(rep(1, 4), rep(2, 4), rep(3, 4), rep(4, 4), rep(5, 4), rep(6, 4)))
```

(a)
We want to use F-test.

$H_0 : \tau_1 = \tau_2 = \tau_3 = \tau_4 = \tau_5 = \tau_6 = 0$, $H_a :$ at least one is not $0$
```{r}
model = lm(y ~ x)
summary(model)
```
From summary, F-statistic is 30.85 with degree of freedoms of 5 and 18. The p-value is $3.16 \times 10^{-8}$. So we have tons of evidence to reject $H_0$. Hence the size of the orifice affects the mean percentage of radon released.

(b)
```{r}
model$residuals
```

```{r}
par(mfrow=c(2,2))
qqnorm(model$residuals)
plot(model$residuals)
plot(model$fitted.values, model$residuals)
```
In Q-Q plot, the residuals lie on a straight line reasonably well. And in the other plots, the residuals lies in a band between 0 and relatively random and there is no obvious patterns.

(c)
The 95% confidence interval for $\hat{\mu} + \tau_5$ is $\hat{\mu} + \hat{\tau_5} \pm c \sqrt{\frac {\hat{\sigma}^2} {2r}}$ where $c \sim t_{18}$.
```{r}
qt(0.975, 18)
```
The CI is,
$72.3750 - 7.3750 \pm 2.1 \frac {2.711} {\sqrt{2 \times 4}} = (62.98719, 67.01281)$

























