---
title: "Untitled"
author: "Mushi Wang 20732874"
date: "10/11/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

4.(a)
```{r}
library(readxl)
chem <- read_excel("clathrates.xls")
fit = lm(chem$y~chem$x1 + chem$x2, chem)
summary(fit)
```
So $y = 11.0870 + 350.1192 x_1 + 0.1089 x_2$

(b)
When there are no surfactant, at the beginning of clathrate formation, 11.087% clathrate has formed.
When the amount of surfactant increased 0.01%, there are 3.5% new formed clathrate compare to the same amount of surfactant at same time.
In 1 mintue, there are 0.1089% new formed clathrate when the amount of surfactant does not change.

(c)
$H_0 : \beta_1 = 0$ vs. $H_a : \beta_1 \neq 0$
form summary(fit), $se(\beta_1) = 39.68$. So $t = \frac {\hat \beta_1 - 0} {se(\beta_1)}$
```{r}
t = as.numeric(fit$coefficients[2] / 39.68)
tdist = qt(0.975, nrow(chem) - 3)
t
tdist
```
Since $\lvert t \rvert > t_{0.975, 33}$ we reject $H_0$. So there is a linear relationship between clathrate formation and amount of surfactant.

$H_0 : \beta_2 = 0$ vs. $H_a : \beta_2 \neq 0$
form summary(fit), $se(\beta_2) = 9.983*10^{-3}$. So $t = \frac {\hat \beta_2 - 0} {se(\beta_2)}$
```{r}
t = as.numeric(fit$coefficients[3] / 0.009983)
tdist = qt(0.975, nrow(chem) - 3)
t
tdist
```
Since $\lvert t \rvert > t_{0.975, 33}$ we reject $H_0$. There is a linear relationship between clathrate formation and time.

