---
title: "a5"
author: "Mushi Wang"
date: "24/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***q1***
```{r}
options(contrasts = c('contr.sum', 'contr.poly'))
y = c(90.4, 90.2, 90.7, 90.6, 90.2, 90.4, 90.1, 90.3, 90.5, 
      90.6, 89.9, 90.1, 90.5, 90.7, 90.8, 90.9, 90.4, 90.1)
x = as.factor(c('00', '00', '01', '01', '02', '02', '10', '10', '11', 
                '11', '12', '12', '20', '20', '21', '21', '22', '22'))
```
(a)
```{r}
model1 = lm(y ~ x)
anova(model1)
```
$H_0: \tau_{11} = \tau_{12} = \cdots = \tau_{33} = 0$ vs $H_a:$ at least one of them is not zero.

From anova table $d = 8$ and $D \sim F_{8, 9}$

From anova table p-value = 0.002638

Hence there is tons of evidence against $H_0$, therefore there is a treatment effect.

(b)
```{r}
temp = as.factor(c(rep(150, 6), rep(160, 6), rep(170, 6)))
psig = as.factor(rep(c(rep(200, 2), rep(215, 2), rep(230, 2)), 3))

model2 = lm(y ~ temp + psig + temp * psig)
anova(model2)
```
$H_0:$ no interaction vs $H_a:$interaction 

From anova table, $d = \frac {ms_{int}} {ms_{res}} = \frac {0.01722} {0.01778} = 0.9685039$, $D \sim F_{4, 9}$

From anova table, p-value = $0.4700058$

There is no evidence against $H_0$, hence no interaction.

(c)
```{r}
#calculate mean for each 
hat = vector(length = 9)
calmean = function(m, v) {
  j = 1
  for (i in seq(from = 1, to = 17, by = 2)) {
    m[j] = (v[i] + v[i + 1]) / 2  
    j = j + 1
  }
  return(m)
}
hat = calmean(hat, y)

plot(x = c(rep(150, 3), rep(160, 3), rep(170, 3)), 
     y = hat,
     xlab = "temp", ylab = "yield")

clrs = c('black', 'red', 'blue')
for (i in 1:3) {
  lines(x = c(150, 160, 170), y = c(hat[i], hat[i + 3], hat[i + 6]), col = clrs[i])
}

legend("topleft",  bty='n', cex=1,
       legend = c("200", "215", "230"),
        col = c('black', 'red', 'blue'),
       lty=c(1, 1), lwd=c(2,2),
       text.width = 1
)
```
From the plot, the groups where pressure is 230 and the groups where pressure is 200 at temperature 150 and 160 have a little interaction. The other lines are roughly parallel to each other. So there is a little interaction.






