---
title: "a68"
author: "Mushi Wang"
date: "15/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.(a)
```{r}
com = read.csv("commissions.csv")
plot(com$Hours_Worked, com$Commission, xlab = "hours worked", ylab = "commission")
```
Since there is a roughly linear relationship. Also the plot is bell shaped, the variance of residual is not a constant. Hence we want to use a ratio estimation.

\newpage
1.(b)
```{r}
mean(com$Commission)
mean(com$Hours_Worked)
length(com$Unit)
306.1579 / 22.15789 * 21
```

$\hat{\mu}_{ratio} = \frac {\bar{y}} {\bar{x}} \mu_x = \frac {306.1579} {22.15789} \cdot 21 = 290.1592$

```{r}
sqrt_hours = sqrt(com$Hours_Worked)
com_ratio = data.frame(unit = com$Unit, hours_worked = com$Hours_Worked / sqrt_hours, 
                       commission = com$Commission / sqrt_hours)
com_ratio_model = lm(com_ratio$commission ~ com_ratio$hours_worked - 1)
summary(com_ratio_model)
```
From summary, $\hat{\sigma}_{ratio} = 29.07$

So, a 95% C.I. for the mean commission is 
\begin{align*}
  & \hat{\mu}_{ratio} \pm \frac {c \hat{\sigma}_{ratio}} {\sqrt n} \sqrt{1 - \frac n N} \text{\quad where $C \sim N(0, 1)$}\\
  = & 290.1592 \pm 1.96 \frac {29.07} {\sqrt 38} \sqrt{1 - \frac {38} {112}}\\
  = & [282.6462, 297.6722]
\end{align*}
We are 95% confident that the mean commission lies in that interval.

Since $\tilde Y_{tot} = N \cdot \tilde \mu_{ratio}$.

$\hat{y}_{tot} = N \cdot \hat{\mu}_{ratio} = 32497.83$, 
$sd(\hat y_{tot}) = \sqrt {N^2 \cdot var(\mu_{ratio})} = N \cdot sd(\mu_{ratio})$

a 95% C.I. for the total commission is 
\begin{align*}
  & \hat{\mu}_{ratio} \pm \frac {c N \hat{\sigma}_{ratio}} {\sqrt n} \sqrt{1 - \frac n N} \text{\quad where $C \sim N(0, 1)$}\\
  = & 32497.83 \pm 1.96 \times 112 \frac {29.07} {\sqrt 38} \sqrt{1 - \frac {38} {112}}\\
  = & [31656.37, 33339.29]
\end{align*}
We are 95% confident that the total commission lies in that interval.

\newpage
1.(c)
$\frac {19} {20} = 0.95$
\begin{align*}
  \frac {c \hat{\sigma}_{ratio}} {\sqrt n} \sqrt{1 - \frac n N} & \leq 1 \text{\quad where $C \sim N(0, 1)$}\\
  \sqrt{\frac 1 n - \frac {1} {112}} & \leq \frac {1} {1.96 \times 29.07}\\
  \frac 1 n & \leq \frac {1} {3246.401} + \frac {1} {112}\\
  n & \geq 108.2649
\end{align*}
So we need at least 109 employers

\newpage
1.(d)
```{r}
sqrt(var(com$Commission))
```
$\hat{\sigma}_{y} = 185.2197$, $\hat{\mu}_y = \bar y = 306.1579$

So, a 95% C.I. is
\begin{align*}
  & \hat{\mu}_{y} \pm \frac {c \hat{\sigma}_{y}} {\sqrt n} \sqrt{1 - \frac n N} \text{\quad where $C \sim N(0, 1)$}\\
  = & 306.1579 \pm 1.96 \frac {185.2197} {\sqrt 38} \sqrt{1 - \frac {38} {112}}\\
  = & [258.2885, 354.0273]
\end{align*}

This C.I. is much wider than the ratio intervals since $\hat \sigma_y$ is much larger. Hence the ratio interval is more accurate.
















